---
title: "Hack Oregon - Exploring Committee Transactions"
author: "John D. Smith and Allan Grimm"
date: "October 6-8, 2015"
output: html_document
---
We had a very open-ended invitation to explore one of several datasets on 

  http://totalgood.github.io/hackor/
  
We chose to explore the Raw Committe Transactions table.

The following describes more or less where our exploration ended up, rather than showing the *many* false starts and dead ends that we encountered.
```{r}
library(dplyr)
library(lubridate)
raw_trans  <- read.delim("/Users/jds/Desktop/hackor/raw_committee_transactions.csv", stringsAsFactors = F, sep = ",")
```
The raw dataset has `r dim(raw_trans)[1]` rows and  `r dim(raw_trans)[2]` columns.

```{r}
#  Correct the data types and add a flag for negative amounts.
raw_trans <- raw_trans %>% 
  mutate(tran_id = as.character(tran_id),
         original_id = as.character(original_id),
         atteset_date = ymd(attest_date),
         review_date = ymd(review_date),
         due_date = ymd(due_date),
         neg_amt_flag = ifelse(amount < 0, "Y", "N")
         )

# have a look at the raw data:
head(raw_trans)

# have a look at negative transactions, because that sound peculiar:
filter(raw_trans, neg_amt_flag == "Y")  %>%  head()

# count by transaction status and neg_amt_flag:
raw_trans %>% group_by(tran_status, neg_amt_flag) %>% summarise(count = n())

# Separate two quite different kinds of records:
am_trans <-  raw_trans %>% filter(tran_status == "Amended")
or_trans <-  raw_trans %>% filter(tran_status == "Original")

last_ammendment <- am_trans %>% group_by(original_id) %>% 
  filter(min_rank(desc(filed_date)) == 1) %>% ungroup()

check_am_trans <- am_trans  %>% filter(original_id == "1555159" | original_id == "1881696" |
                        tran_id == "1555159" | tran_id == "1881696") %>% arrange(original_id)
select(check_am_trans, tran_id, original_id, attest_date, filed_date)

check_last_ammendment <- last_ammendment %>% filter(original_id == "1555159" | original_id == "1881696" |
                        tran_id == "1555159" | tran_id == "1881696") %>% arrange(original_id)

select(check_last_ammendment, tran_id, original_id, attest_date, filed_date)
# conculsion: indeedd the correct record was selected.
```
It's still puzzling why "Ammended" records all point to an "original" but the original transaction ID doesn't seem appear in the dataset.

Because some transactions have been ammended more than once, the filter on the "last transaction" reduces the total number of ammended transactions by `r dim(am_trans)[1] - dim(last_ammendment)[1]` to `r dim(last_ammendment)[1]`.  Should the rows with the "last ammended transaction" be added back to the *un-ammended* transactions?  Probably so.

```{r}

subsequently_ammended <- am_trans %>% group_by(original_id) %>% 
  filter(min_rank(desc(filed_date)) != 1) %>% ungroup()

# check whether multiple ammendments are for negative amounts and count up the total:
subsequently_ammended %>% group_by(neg_amt_flag) %>%  summarise(total = sum(amount))

x <- subsequently_ammended %>%  summarise(total = sum(amount))
```
The original dataset has `r dim(subsequently_ammended)[1]` extra records that should be eliminated as suggested in this exploration. Without filtering out the duplicated adjusted records, total amounts are overstated by $ `r prettyNum(x$total, big.mark = ",", format = "d")`.

```{r}

```
